PREVIC: An adaptive parent report measure of expressive vocabulary in children between 3 and 8 years of age

### 2023-10-16 paper/manuscript.Rmd
prep repo for submission

  - name          : "Jonas Engicht"
    affiliation   : "3"
    role:
      - Formal Analysis
      - Software
      - Writing - Original Draft Preparation
      - Writing - Review & Editing

@@

affiliation:
  - id            : "1"
    institution   : "Institute for Psychology, Leuphana University Lüneburg, Germany"
  - id            : "2"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"
  - id            : "3"
    institution   : "Institut of Psychology, Friedrich-Schiller-University Jena, Germany"
  - id            : "4"
    institution   : "Leibniz-Zentrum Allgemeine Sprachwissenschaft, Berlin, Germany"

@@

abstract: |
  Parent report measures have proven to be a valuable research tool to study early language development. Caregivers are given a list of words and are asked which of them their child has already used. However, most available measures are not suited for children beyond infancy, come with substantial licensing costs or lack a clear psychometric foundation. Here we present the PREVIC (Parent Report of Expressive Vocabulary in Children) an open access, high quality vocabulary checklist for German-speaking children between three and eight years of age. The PREVIC was constructed leveraging the advantages of Item Response Theory: we designed a large initial item pool of 379 words  and collected data from N = 1190 caregivers of children between three and eight years of age. Based on this data, we computed a range of fit indices for each item (word) and used an automated item selection algorithm to compile a final pool that contains items that a) vary in difficulty and b) fit the Rasch (one-parameter logistic) model. The resulting task is highly reliable and shows convergent validity. The IRT-based construction allowed us to design an adaptive version of the task, which substantially reduces the duration of the task while retaining measurement precision. The task -- including the adaptive version -- was implemented as a website and is freely accessible online at https://ccp-odc.eva.mpg.de/previc-demo/. The PREVIC can easily be adapted to other languages and fills an important gap in the toolkit of researchers interested in language development.

keywords          : "language development, vocabulary, individual differences, Item Response Models"
wordcount         : "X"

@@

Child language measures can be broadly categorized into two types: direct and parent report measures. Direct measures are generally used with children of three years and older. Direct expressive language assessments involve prompting children to generate words or sentences in response to a stimulus, such as a picture or an object. Direct receptive language assessments reverse the logic and require children to match a verbal prompt with a picture or an object. Various direct measures tailored to different languages and age groups have been developed, including measures for English and German [@dunn1965peabody; @dunn1997british; @golinkoff2017user; @kiese2005awst; @kauschke2002patholinguistische; @lenhard2015peabody; @gluck2011wortschatz; @bohn2023orev]. Additionally, standardized cognitive ability tests frequently incorporate direct language measures [e.g., @bayley2006bayley; @wechsler1949wechsler; @gershon2013iv].

Parent report measures in general are widely utilized in psychological research. They are particularly popular as screening methods to identify developmental delays [@diamond1993role; @pontoppidan2017parent]. However, it is important to acknowledge that parent reports come with certain caveats, including the potential for selective reporting and social desirability. As a consequence, providing a comprehensive assessment of the overall quality and usefulness of these measures is challenging [@morsbach2006understanding]. Nonetheless, some parent report measures have been found to be both reliable and valid [@hornman2013validity; @ireton1995assessin; @macy2012evidence; @bodnarchuk2004can; @saudino1998validity].

In child language research, parent report measures are often utilized with very young children when direct assessment is challenging. One widely used measure is the MacArthur-Bates Communicative Development Inventories [CDI, @fenson2007macarthur]. The CDI asks parents to to check those words from a checklist that they believe their child produces and/or understands. This measure has been adapted for a wide range of spoken and signed languages [see @frank2021variability for an overview], with various versions available [e.g., @makransky2016item; @mayor2019short], including an online version [@demayo2021web]. Collaborative efforts have facilitated the pooling of CDI data from thousands of children learning different languages into centralized repositories [@frank2017wordbank; @jorgensen2010clex]. Importantly, the CDI exhibits validity as parental reports align with direct observations and assessments of child language [@bornstein1998vocabulary; @fenson1994variability; @dale1991validity; @feldman2005concurrent].

@@

A more general issue with existing language measures -- including PPVT and DVAP -- is a lack of psychometric grounding. Items that make up the scale are selected based on researchers intuitions and there is no clear measurement model that explicates how the different items and test scores are linked to the construct in question [@borsboom2006attack]. Item response theory (IRT) provides a theoretical framework and a toolkit to develop tasks with a solid psychometric foundation [@lord2012applications; @kubinger2006psychologische]. Within IRT, all items are assumed to measure the same latent construct. Each item is linked to the construct by a probability function (e.g. a logistic curve) which determines how likely a particular response is for individuals with different values on the latent construct. The shape of this curve is defined by the difficulty of an item (i.e. the value of the latent construct when the probability to solve the item is 50%) and its discrimination (i.e. the slope of the curve showing how the probability to solve the item changes with increasing levels on the latent construct). In a *Rasch* model (or one-parameter IRT model), all items are assumed to have equal item discriminations, resulting in parallel item characteristic curves. The great benefit of IRT is that models are testable in that we can quantify the fit of each model and compare competing models. For each item, we can compute fit statistics that indicate how well the model captures the response pattern to the item. Test construction is straightforward in this framework; items are selected that improve the fit to the model. Besides many other advantages (e.g., objective specificity, sum scores can be used as sufficient statistics), the Rasch model also allows for adaptive testing. Here, individuals do not simply respond to all items in the task but only to a subset of optimally informative items given their -- constantly updated -- value on the latent construct.

The downside of IRT-based test construction -- and probably the reason it is not used more often -- is that it requires a larger initial investment. To be able to remove items with a poor fit during the selection process requires an initial item pool that is substantially larger than the desired size of the final task. Adaptive testing also needs a large item pool so that there is a sufficient number of items that are optimally informative for different regions of the latent construct. Furthermore, to obtain solid estimates for the item parameters it takes large sample sizes. Yet, we believe that these disadvantages are clearly outweighed by the benefits that come with IRT-based test construction in the long run.

@@

We decided to use an interactive format instead of presenting parents with a long list of words. The task was implemented as a web-app using `html` and `JavaScript` and ran in every modern web-browser on computers, tablets and smartphones. Words were presented one-by-one and caregivers could indicate whether or not their child says a word either by using the familiar swipe-left/swipe-right functionality, by clicking symbols or using arrow-keys on a keyboard (see Figure \@ref(fig:fig1)A). For example, caregivers saw the word "Jacke" (en: jacket) on a card in the center of the screen; to report that their child says the word, they would swipe the card to the right side of the screen which would make the card go away and the next in the deck appear. We included a lightweight back-end that registered the last completed trial so that caregivers could take breaks and even switch devices during the task.

@@

Our goal was to create an item pool with items of varying difficulty. We used age-of-acquisition (AoA) ratings as a rough indicator of anticipated item difficulty. Previous work has shown strong associations between AoA ratings and how likely children are to know a word [@bohn2023orev; @bohn2021young]. We started the process by compiling a list with AoA ratings for 3,921 German words from various sources [@luniewska2019age; @schroder2012german; @birchenough2017rated]. We excluded words with AoA ratings above ten. The remaining words were ordered by rated AoA and then split into ten lists with 344 words each. A research assistant with a background in linguistics went through the lists and selected words that a) were indicative of language abilities more broadly (avoiding very specialized terms) and b) that were different from one another in that they were semantically unrelated (to avoid words that are learned in the same context). For each list, we aimed for roughly 35 words; 17 nouns, nine verbs and nine adjectives. The so-generated item pool had `r data%>%distinct(word)%>%summarise(n())` words, of which `r words$noun` were nouns, `r words$verb` were verbs and `r words$adjective` were adjectives. Figure \@ref(fig:fig2)A shows how the items were distributed across AoA ratings and word types.

@@

The goal of the item selection procedure was to generate an item pool with items that fit the Rasch model. We selected items in three steps. In the first step, we excluded items using conventional cut-offs for indices that quantify the fir of each item to the Rasch model. The goal was to remove a large number of items with a poor fit to reduce the computational burden in subsequent steps. In step 2, we used an automated item selection procedure to select an optimal subset of items from the remaining pool. We focused on the fit of the Rasch model as well as variation in item difficulty (to measure in different regions of the latent dimension). Finally, in step 3, we submitted the items selected in step 2 to an analysis of differential item functioning (DIF).

IRT-models were implemented in a Bayesian framework in `R` using the `brms` package [@burkner2019bayesian; @burkner2017brms] unless otherwise stated. We predicted the probability of a correct answer based on a participant's latent ability and item characteristics. We fit two classes of models: Rasch (1PL) Birnbaum (2PL) models. The main difference between these two models lies in their assumption about how the probability of solving an item changes with ability levels (item discrimination). Here, the Rasch model assumes that the rate of change (i.e. the slope of the logistic curve) is the same for all items while the 2PL model allows item discrimination parameters to vary between items.

@@

We fit a Rasch Model to the data and computed In- and Outfit values based on draws from the model's posterior predictive distribution (using the function `add_epred_draws`) for each item and person combination. In- and Outfit quantify the deviation of a person's response to an item from what the model predicts based on the difficulty of an item and the person's ability parameter. The result was a distribution of values for each item and index. For each item, we then computed the mode for each index. The closer to 1 the index is, the better the fit. Figure \@ref(fig:fig3) visualizes the results. We used the cut-off values suggested in the literature [@bond2013applying; @debelak2022introduction] and excluded items with In- or Outfit values below 0.7 and above 1.3. Like all heuristics, these cut-offs are to some extent arbitrary. Yet, in the present context they served the purpose of removing a large number of potentially unsuitable items. This procedure led us to exclude `r fit_select$exclude` of the `r n_distinct(data$word)` items in the pool, leaving `r fit_select$include` for the automated item selection.

@@

The goal of this step was to select items with different levels of difficulty that fit the Rasch model. Selecting items based on these criteria ensured that a) the final item pool allowed for precise measurement in different regions of the latent ability and b) the number of solved items is a sufficient statistic for an individual's ability. Such an item pool is then optimally suited for adaptive testing.

@@

Figure \@ref(fig:fig3)B visualizes the model comparison and shows that the Rasch model provided a good fit for subsets up to 90 items. We therefore decided on 90 items as the size of the final item pool. We ran the simulated annealing algorithm 20 times and selected the 90 items that were returned most often (86 items were returned on every run).

@@

link_data <- data%>%
  filter(word %in% final_items)%>%
  group_by(subjID)%>%
  summarise(previc = sum(score))%>%
  left_join(orev_data%>%
              group_by(subjID)%>%
              summarise(orev = sum(correct)))%>%
  filter(!is.na(orev))%>%
  ungroup()%>%
  distinct(subjID, .keep_all = T)

@@

We found a substantial correlation between caregiver's answers to questions about their children's expressive vocabulary in the PREVIC and a direct assessment of children's receptive vocabulary in the oREV (*r* = `r round(val_cor$estimate,2)`; 95% CI = `r round(val_cor$conf.int[1],2)` -- `r round(val_cor$conf.int[2],2)`; Figure \@ref(fig:fig4)B)). Furthermore, when predicting PREVIC scores by oREV scores and age in a binomial model, we found that oREV scores had a large and positive effect (Figure \@ref(fig:fig4)C). These results speak to the convergent validity of the PREVIC.

(ref:figlab4) Item characteristics and validity. (A) Item characteristic curves for the 89 items colored by word type. Inset on the upper left shows the test information curve. (B) Correlation between PREVIC and oREV scores.Points show aggregated scores of individuals in the two tasks. Points have minimal horizontal noise added to avoid overplotting. The red line shows a regression line (with 95% CI) based on a linear model. (C) Posterior model estimates for oREV scores and age (scaled) in a model predicting PREVIC scores. Points show posterior means with 95% CrI.

@@

The large and diverse item pool allowed us to to create an adaptive version of the PREVIC in addition to the complete checklist. The general idea of an adaptive test is to only show the caregiver the most informative items given the (continuously updated) ability. As a consequence, items that are too easy or too difficult are omitted and the test becomes substantially shorter while retaining the same level of measurement precision.

In order to determine the most informative items, the ability of the child has to be estimated during the test. To achieve this, we implemented a maximum likelihood estimator in `html` and `TypeScript` (which is compiled to native `JavaScript`). As a consequence, the adaptive version is still fully portable and can be run in any modern web-browser. The estimated ability "is the ability value that maximizes the likelihood function $L(\theta)$" [@magis2012] given the item response $y_i$ (either 0 or 1) and the item difficulty $\alpha_i$ [@eid2011; @eid2014testtheorie]:

@@

At the beginning of the test, the ability level is set to 0. A person specific ability estimation is also not possible using Maximum Likelihood after the first item and we therefore followed the convention to set the ability level to -4 if the answer was "no" and 4 if the answer was "yes" [e.g. implemented in the `R` package `catR`, @catR]. The test then continues until a pre-specified level of measurement precision (standard error of the ability estimate) is reached or until all items have been used. Users can set the desired level of measurement precision at the beginning of the test (e.g. SE of 0.3, 0.4 or 0.5), which again influences its length (larger SE means shorter test). In the end, the user downloads a file containing the final ability level, the SE at which the test stopped, the answered items (including the word itself and its difficulty) and the participant's response pattern.

We validated the implementation of our estimator by comparing its ability estimates and selected items to those of the `catR` package in a number of simulations. The results were identical and only differed beyond the fifth decimal because `JavaScript` and `R` differ in how many decimal places they use for computations. The code to run the simulations can be found in the associated online repository.

@@

This paper describes the construction and validation of the PREVIC, an adaptive parent report measure of productive vocabulary in German-speaking children between three and eight years of age. Following the logic of widely-used vocabulary checklists for younger children [@fenson2007macarthur], the PREVIC presents caregivers with individual words and asks if the child speaks this word. The items (words) that make up the PREVIC were selected using Item-Response theory (IRT): we started with a large initial item pool of 379 words from which we selected 89 items that fit the Rasch model. The resulting task is highly reliable and shows convergent validity when contrasted with a direct receptive vocabulary measure. Leveraging IRT allowed us to devise an adaptive version of the task in which only the most informative items are presented. The task is implemented as a web-app and can be used with any device that runs a modern web browser. The task itself (adaptive and complete checklist) as well as the source code are freely available online.

The PREVIC fills an important gap in the tool kit of researchers studying language development beyond infancy. It complements direct assessments of children's vocabulary by providing an additional perspective on children's vocabulary skills. Parents observe children for extended periods of time and their assessment therefore provides a more aggregated measure. Parent reports are also immune to momentary fluctuations in children's motivation and attention that might influence the results of direct assessments. Nevertheless, parental reports remain indirect measures and are ideally combined with direct assessments whenever feasible. Given that the PREVIC is short -- in particular the adaptive version rarely takes more than five minutes to complete -- it can easily be filled out by parents during a lab visit. Its implementation as a web-app even allows for sending it to families before or after a visit.

At present, the PREVIC is only available in German. However, with inclusivity and broad applicability in mind, we have made the entire source code available. This not only facilitates its adaptation to other languages and allows researchers to use the same user interface. Encouragingly, preliminary feedback indicates that parents find the interface intuitive and user-friendly. The CDI has seen expansive adaptation across various languages [see @frank2021variability for a summary]. Such adaptations are usually not complete translations in that some words are removed and others added to capture the linguistic nuances and specificities of each language. For the adaptation of the PREVIC to new languages, a comprehensive reassessment of item properties would be highly desirable. For adaptive testing, it would even be mandatory. Taken together, we hope that our commitment to openness will put the PREVIC on a similar trajectory as the CDI.

@@

@@

The task can be accessed via the following website: <https://ccp-odc.eva.mpg.de/previc/>. The corresponding source code can be found in the following repository: <https://github.com/ccp-eva/previc-demo>. The data sets generated during and/or analysed during the current study are available in the following repository: <https://github.com/ccp-eva/previc-analysis/>. Data collection was preregistered at: <https://osf.io/utzfh>.

>>>>

  - name          : "Jonas Engicht"
    affiliation   : "3"
    role:
      - Software
      - Writing - Original Draft Preparation
      - Writing - Review & Editing

@@

affiliation:
  - id            : "1"
    institution   : "Institute for Psychology, Leuphana University Lüneburg, Germany"
  - id            : "2"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"
  - id            : "3"
    institution   : "Institute of Psychology, Friedrich-Schiller-University Jena, Germany"
  - id            : "4"
    institution   : "Leibniz-Zentrum Allgemeine Sprachwissenschaft, Berlin, Germany"

@@

abstract: |
  Parent report measures have proven to be a valuable research tool to study early language development. Caregivers are given a list of words and are asked which of them their child has already used. However, most available measures are not suited for children beyond infancy, come with substantial licensing costs or lack a clear psychometric foundation. Here we present the PREVIC (Parent Report of Expressive Vocabulary in Children), an open access, high quality vocabulary checklist for German-speaking children between three and eight years of age. The PREVIC was constructed leveraging the advantages of Item Response Theory: we designed a large initial item pool of 379 words and collected data from N = 1190 caregivers of children between three and eight years of age. Based on this data, we computed a range of fit indices for each item (word) and used an automated item selection algorithm to compile a final pool that contains items that a) vary in difficulty and b) fit the Rasch (one-parameter logistic) model. The resulting task is highly reliable and shows convergent validity. The IRT-based construction allowed us to design an adaptive version of the task, which substantially reduces the duration of the task while retaining measurement precision. The task -- including the adaptive version -- was implemented as a website and is freely accessible online (https://ccp-odc.eva.mpg.de/previc-demo/). The PREVIC fills an important gap in the toolkit of researchers interested in language development and provides an ideal starting point for the development of converging measures in other languages.

keywords          : "language development, vocabulary, individual differences, Item Response Models"
wordcount         : "5711"

@@

Child language measures can be broadly categorized into two types: direct and parent report measures. Direct measures of productive and receptive language are generally used with children of three years and older. Direct expressive language assessments involve prompting children to generate words or sentences in response to a stimulus, such as a picture or an object. Direct receptive language assessments reverse the logic and require children to match a verbal prompt with a picture or an object. Various direct measures tailored to different languages and age groups have been developed, including measures for English and German [@dunn1965peabody; @dunn1997british; @golinkoff2017user; @kiese2005awst; @kauschke2002patholinguistische; @lenhard2015peabody; @gluck2011wortschatz; @bohn2023orev; @armon2015assessing]. Additionally, standardized cognitive ability tests frequently incorporate direct language measures [e.g., @bayley2006bayley; @wechsler1949wechsler; @gershon2013iv].

Parent report measures in general are widely utilized in psychological research. They are particularly popular as screening methods to identify developmental delays [@diamond1993role; @pontoppidan2017parent]. However, it is important to acknowledge that parent reports come with certain caveats, including the potential for selective reporting and social desirability. As a consequence, providing a comprehensive assessment of the overall quality and usefulness of these measures is challenging [@morsbach2006understanding]. Nonetheless, some parent report measures have been found to be both reliable and valid [@hornman2013validity; @ireton1995assessin; @macy2012evidence; @bodnarchuk2004can; @saudino1998validity; @de2022quantifying].

In child language research, parent report measures are often utilized with very young children when direct assessment is challenging. One widely used measure is the MacArthur-Bates Communicative Development Inventories [CDI, @fenson2007macarthur]. The CDI asks parents to check those words from a checklist that they believe their child produces and/or understands. This measure has been adapted for a wide range of spoken and signed languages [see @frank2021variability for an overview], with various versions available [e.g., @makransky2016item; @mayor2019short], including an online version [@demayo2021web]. Collaborative efforts have facilitated the pooling of CDI data from thousands of children learning different languages into centralized repositories [@frank2017wordbank; @jorgensen2010clex]. Importantly, the CDI exhibits validity as parental reports align with direct observations and assessments of child language [@bornstein1998vocabulary; @fenson1994variability; @dale1991validity; @feldman2005concurrent].

@@

A more general issue with existing language measures -- including PPVT and DVAP -- is a lack of psychometric grounding. Items that make up the scale are selected based on researchers intuitions and there is no clear measurement model that explicates how the different items and test scores are linked to the construct in question [@borsboom2006attack]. Item response theory (IRT) offers a theoretical framework to fill this gap and provides a toolkit to develop tasks with a solid psychometric foundation [@lord2012applications; @kubinger2006psychologische]. Within IRT, all items are assumed to measure the same latent construct. Each item is linked to the construct by a probability function (e.g. a logistic curve) which determines how likely a particular response is for individuals with different values on the latent construct. The shape of this curve is defined by the difficulty of an item (i.e. the value of the latent construct when the probability to solve the item is 50%) and its discrimination (i.e. the slope of the curve showing how the probability to solve the item changes with increasing levels on the latent construct). In a *Rasch* model (or one-parameter IRT model), all items are assumed to have equal item discriminations, resulting in parallel item characteristic curves. The great benefit of IRT is that models are testable in that we can quantify the fit of each model and compare competing models. For each item, we can compute fit statistics that indicate how well the model captures the response pattern to the item. Test construction is straightforward in this framework; items are selected that improve the fit to the model.

Besides many other advantages (e.g., objective specificity, sum scores can be used as sufficient statistics), the Rasch model also allows for adaptive testing. All items in a test that conforms to the Rasch model measure the same latent ability and all that differs is their difficulty. As a consequence, the ability of a person can be estimated independent of the particular items that have been completed. This characteristic is leveraged during adaptive testing: individuals do not simply respond to all items in the task but only to the items that are optimally informative given their -- constantly updated -- individual value on the latent construct. Because all items measure the same latent ability, the resulting scores are nevertheless comparable.

The downside of IRT-based test construction -- and probably the reason it is not used more often -- is that it requires a larger initial investment. To be able to remove items with a poor fit during the selection process requires an initial item pool that is substantially larger than the desired size of the final task. Adaptive testing also needs a large item pool so that there is a sufficient number of items that are optimally informative for different regions of the latent construct. Furthermore, to obtain solid estimates for the item parameters it takes large sample sizes. Yet, we believe that these initial costs are clearly outweighed by the benefits that come with IRT-based test construction in the long run.

@@

We decided to use an interactive format instead of presenting parents with a long list of words in order to increase number of items while keeping the task engaging. The task was implemented as a web-app using `html` and `JavaScript` and ran in every modern web-browser on computers, tablets and smartphones. Words were presented one-by-one and caregivers could indicate whether or not their child says a word either by using the familiar swipe-left/swipe-right functionality, by clicking symbols or using arrow-keys on a keyboard (see Figure \@ref(fig:fig1)A). For example, caregivers saw the word "Jacke" (en: jacket) on a card (color-coded by part of speech: noun = blue, adjective = orange, verb = green) in the center of the screen; to report that their child says the word, they would swipe the card to the right side of the screen which would make the card go away and the next in the deck appear. We included a lightweight back-end that registered the last completed trial so that caregivers could take breaks and even switch devices during the task.

@@

Our goal was to create an item pool with items of different word classes and varying semantical difficulty. We used age-of-acquisition (AoA) ratings as a rough indicator of anticipated item difficulty. Previous work has shown strong associations between AoA ratings and how likely children are to know a word [@bohn2023orev; @bohn2021young]. We started the process by compiling a list with AoA ratings for 3,921 German words from various sources [@luniewska2019age; @schroder2012german; @birchenough2017rated]. We excluded words with AoA ratings above ten. The remaining words were ordered by rated AoA and then split into ten lists with 344 words each. A research assistant with a background in linguistics went through the lists and selected words that a) were indicative of language abilities more broadly (avoiding very specialized terms) and b) that were different from one another in that they were semantically unrelated (to avoid words that are learned in the same context). For each list, we aimed for roughly 35 words from the three word classes; 17 nouns, nine verbs and nine adjectives. The so-generated item pool had `r data%>%distinct(word)%>%summarise(n())` words, of which `r words$noun` were nouns, `r words$verb` were verbs and `r words$adjective` were adjectives. Figure \@ref(fig:fig2)A shows how the items were distributed across AoA ratings and word types.

@@

The goal of the item selection procedure was to generate an item pool with items that fit the Rasch model. We selected items in three steps. In the first step, we excluded items using conventional cut-offs for indices that quantify the fit of each item to the Rasch model. The goal was to remove a large number of items with a poor fit to reduce the computational burden in subsequent steps. In step 2, we used an automated item selection procedure to select an optimal subset of items from the remaining pool. We focused on the fit of the Rasch model as well as variation in item difficulty (to measure in different regions of the latent dimension). Finally, in step 3, we submitted the items selected in step 2 to an analysis of differential item functioning (DIF).

IRT-models were implemented in a Bayesian framework in `R` using the `brms` package [@burkner2019bayesian; @burkner2017brms] unless otherwise stated. We predicted the probability of a correct answer based on a participant's latent ability and item characteristics. We fit two classes of models: Rasch (1PL) and Birnbaum (2PL) models. The main difference between these two models lies in their assumption about how the probability of solving an item changes with ability levels (item discrimination). Here, the Rasch model assumes that the rate of change (i.e. the slope of the logistic curve) is the same for all items while the 2PL model allows item discrimination parameters to vary between items.

@@

We fit a Rasch Model to the data and computed In- and Outfit values based on draws from the model's expected posterior predictive distribution (using the function `add_epred_draws`) for each item and person combination. In- and Outfit quantify the deviation of a person's response to an item from what the model predicts based on the difficulty of an item and the person's ability parameter. The result was a distribution of values for each item and index. For each item, we then computed the mode for each index. The closer to 1 the index is, the better the fit. Figure \@ref(fig:fig3) visualizes the results. We used the cut-off values suggested in the literature [@bond2013applying; @debelak2022introduction] and excluded items with In- or Outfit values below 0.7 and above 1.3. Like all heuristics, these cut-offs are to some extent arbitrary. Yet, in the present context they served the purpose of removing a large number of potentially unsuitable items. This procedure led us to exclude `r fit_select$exclude` of the `r n_distinct(data$word)` items in the pool, leaving `r fit_select$include` for the automated item selection.

@@

The goal of this step was to select items with different levels of difficulty that fit the Rasch model. Selecting items based on these criteria ensured that a) the final item pool allowed for precise measurement in different regions of the latent ability and b) the number of solved items is a sufficient statistic for an individual's ability. Such an item pool is then optimally suited for adaptive testing because items differ in difficulty but measure the same latent dimension. This way, individuals with different ability levels can be shown different items while still ensuring that the eventual scores are directly comparable.

@@

Figure \@ref(fig:fig3)B visualizes the model comparison and shows that the Rasch model provided a good fit for subsets up to 90 items. We therefore decided on 90 items as the size of the final item pool. We ran the simulated annealing algorithm 20 times and selected the 90 items that were returned most often (the same 86 items were returned on every run).

@@

link_data <- data%>%
  filter(word %in% readRDS( "../saves/final_items.rds"))%>%
  group_by(subjID)%>%
  summarise(previc = sum(score))%>%
  left_join(read_csv("../data/orev_data.csv")%>%
              group_by(subjID)%>%
              summarise(orev = sum(correct)))%>%
  filter(!is.na(orev))%>%
  ungroup()%>%
  distinct(subjID, .keep_all = T)

@@

We found a substantial correlation between caregiver's answers to questions about their children's expressive vocabulary in the PREVIC and a direct assessment of children's receptive vocabulary in the oREV (*r* = `r round(val_cor$estimate,2)`; 95% CI = `r round(val_cor$conf.int[1],2)` -- `r round(val_cor$conf.int[2],2)`; Figure \@ref(fig:fig4)B). Furthermore, when predicting PREVIC scores by oREV scores and age in a binomial model, we found that oREV scores had a large and positive effect (Figure \@ref(fig:fig4)C). These results speak to the convergent validity of the PREVIC.

(ref:figlab4) Item characteristics and validity. (A) Item characteristic curves for the 89 items colored by word type. Inset on the upper left shows the test information curve. (B) Correlation between PREVIC and oREV scores. Points show aggregated scores of individuals in the two tasks. Points have minimal horizontal noise added to avoid overplotting. The red line shows a regression line (with 95% CI) based on a linear model. (C) Posterior model estimates for oREV scores and age (scaled) in a model predicting PREVIC scores. Points show posterior means with 95% CrI.

@@

The large and diverse item pool allowed us to to create an adaptive version of the PREVIC in addition to the complete checklist. The general idea of an adaptive test is to only show the caregiver the most informative items given the (continuously updated) individual ability. As a consequence, items that are too easy or too difficult are omitted and the test becomes substantially shorter while retaining the same level of measurement precision.

In order to determine the most informative items, the ability of the child has to be estimated during the test. To achieve this, we implemented a maximum likelihood estimator in `html` and `TypeScript` (which is compiled to native `JavaScript`). As a consequence, the adaptive version is still fully portable and can be run in any modern web-browser. The estimated ability "is the ability value that maximizes the likelihood function $L(\theta)$" [@magis2012], given the item response $y_i$ (either 0 or 1) and the item difficulty $\alpha_i$ [@eid2011; @eid2014testtheorie]:

@@

At the beginning of the test, the ability level is set to 0. A person-specific ability estimation is not yet possible using Maximum Likelihood Estimation after the first item and we therefore followed the convention to set the ability level to -10 if the answer was "no" and 10 if the answer was "yes" [e.g. implemented in the `R` package `catR`, @catR]. The test then continues until a pre-specified level of measurement precision (standard error of the ability estimate) is reached or until all items have been used. Users can set the desired level of measurement precision at the beginning of the test (e.g. SE of 0.3, 0.4 or 0.5), which again influences its length (larger SE means shorter test). In the end, the user downloads a file containing the final ability level, the SE at which the test stopped, the answered items (including the word itself and its difficulty) and the participant's response pattern.

We validated the implementation of our estimator by comparing its ability estimates and selected items to those of the `catR` package in a number of simulations. The results were identical and only differed beyond the fifth decimal because `JavaScript` and `R` differ in their implemented floating-point number format. The code to run the simulations can be found in the associated online repository.

@@

This paper describes the construction and validation of the PREVIC, an adaptive parent report measure of productive vocabulary in German-speaking children between three and eight years of age. Following the logic of widely-used vocabulary checklists for younger children [@fenson2007macarthur], the PREVIC presents caregivers with individual words and asks if the child speaks this word. The items (words) that make up the PREVIC were selected using Item-Response theory: we started with a large initial item pool of 379 words from which we selected 89 items that fit the Rasch model. The resulting task is highly reliable and shows convergent validity when contrasted with a direct receptive vocabulary measure. Leveraging IRT allowed us to devise an adaptive version of the task in which only the most informative items are presented. The task is implemented as a web-app and can be used with any device that runs a modern web browser. The task itself (adaptive and complete checklist) as well as the source code are freely available online.

The PREVIC fills an important gap in the tool kit of researchers studying language development beyond infancy and particular during the preschool years. It complements direct assessments of children's vocabulary by providing an additional perspective on children's vocabulary skills. Parents observe children for extended periods of time and their assessment therefore provides a more aggregated measure. Parental reports are also immune to momentary fluctuations in children's motivation and attention that might influence the results of direct assessments. Nevertheless, parental reports remain indirect measures and are ideally combined with direct assessments whenever feasible. Given that the PREVIC is short -- in particular the adaptive version rarely takes more than five minutes to complete -- it can easily be filled out by parents during a lab visit. Its implementation as a web-app even allows for sending it to families before or after a visit.

At present, the PREVIC is available only in German. However, with inclusivity and broad applicability in mind, we have made the entire source code available. This not only facilitates its adaptation to other languages and allows researchers to use the same user interface. Encouragingly, preliminary feedback indicates that parents find the interface intuitive and user-friendly. The CDI has seen expansive adaptation across various languages [see @frank2021variability for a summary]. Such adaptations are usually not complete translations in that some words are removed and others added to capture the linguistic nuances and specificities of each language. However, @luniewska2019age found that the order of acquisition (and thus presumably the difficulty) was similar for many words across seven languages. Hence, most items in the current pool could be translated and re-used if the PREVIC were to adapted to different languages. Nevertheless, a comprehensive reassessment of item properties would be highly desirable. For adaptive testing, it would even be mandatory. Taken together, we hope that our commitment to openness will put the PREVIC on a similar trajectory as the CDI.

@@

Many points of criticism that apply to parental report measures apply to the PREVIC as well. Parents might be biased in their assessment and more recent events might have a stronger influence on their responses compared to more distant ones. Furthermore, compared to measures for younger children, the PREVIC might be less accurate in and absolute sense because children speak much more words and parents spent less time with their children as they get older. Nevertheless, we found a substantial correlation with a direct measure suggesting that the PREVIC accurately captures relative individual differences.

@@

The task can be accessed via the following website: <https://ccp-odc.eva.mpg.de/previc-demo/>. The corresponding source code can be found in the following repository: <https://github.com/ccp-eva/previc-demo>. The data sets generated during and/or analysed during the current study are available in the following repository: <https://github.com/ccp-eva/previc/>. Data collection was preregistered at: <https://osf.io/utzfh>.
